---
title: "SEM_2_-_Multifactor_models"
author: "Aleksander Molak"
date: "February 5, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
path = "C:\\Users\\aleksander.molak\\Documents\\EDU\\DataCamp_-_StructEqModelling\\Struct_Eq_Modelling"
data_file = "ESP_DATA6_HOMEWORK1.SAV"
color  = '#195e8c'
color2 = '#d4ff00'
color3 = '#b80031'
color4 = '#ffbf00'
```

```{r}
library(lavaan)
library(psych)
library(GGally)

```

# Structural Equation Modeling Exercises #2

## Get data 

```{r}
data(HolzingerSwineford1939)

holzinger <- HolzingerSwineford1939
```


## Define & fit a two-factor model

```{r}
# Define
twofactor_model_def <- 'visual =~ x1 + x2 + x3
                        speed  =~ x7 + x8 + x9'
```


```{r}
# Fit
twofactor_model_fit <- cfa(model = twofactor_model_def,
                           data  = HolzingerSwineford1939)

```

```{r}
summary(twofactor_model_fit, standardized = TRUE, fit.measures = TRUE)
```


Note, that in the two-factor model the output contains `~~`. It signifies covariance.


It's a great moment to introduce a bit more of syntax for model specificatation:


  * `=~` creates latent variables
  * `~~` creates covariance between variables
  * `~` creates direct prediction between variables

Remember, `y ~ x` to specify direction


## Improving model fit - modification indices


* Compare data variance with model variance

```{r}
var(holzinger$x1)
```

We can see that `x1` variance is higher than model-estimated variance for this predictor. This may indicate model variance issue.


* Modification indices

Let's compute modification indices. They'll potentially allow us to find potential parameters (paths) to add to the model specification to improve model fit. 


```{r}
modificationindices(twofactor_model_fit, sort = TRUE)
```


`x7 ~~ x8` has the highest `mi`. Let's add it to the model.

```{r}
improved_model_def <- 'visual =~ x1 + x2 + x3
                       speed  =~ x7 + x8 + x9
                       x7 ~~ x8'
```

```{r}
# Fit
improved_model_fit <- cfa(model = improved_model_def,
                          data  = HolzingerSwineford1939)
```

```{r}
summary(improved_model_fit, standardized = TRUE, fit.measures = TRUE)
```

`TLI` and `CFI` are now much higher and `RMSEA` is around twice lower. Good job! 

But... let's compare model fits statistically:

```{r}
anova(twofactor_model_fit, improved_model_fit)
```


```{r}
fitmeasures(twofactor_model_fit, c("aic", "ecvi"))
```

```{r}
fitmeasures(improved_model_fit, c("aic", "ecvi"))
```


It looks like the improved model is better in terms of $\chi^2$, `AIC` and `ecvi`.


