---
title: 'SEM #1'
author: "Aleksander Molak"
date: "February 3, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
path = "C:\\Users\\aleksander.molak\\Documents\\EDU\\DataCamp_-_StructEqModelling\\Struct_Eq_Modelling"
data_file = "ESP_DATA6_HOMEWORK1.SAV"
color  = '#195e8c'
color2 = '#d4ff00'
color3 = '#b80031'
color4 = '#ffbf00'
```

```{r}
library(lavaan)
library(psych)
library(GGally)

```

# Structural Equation Modeling Exercises #1

## Get data 

```{r}
data(HolzingerSwineford1939)

holzinger <- HolzingerSwineford1939
```


## EDA

```{r}
summary(holzinger)
```

```{r}
ggpairs(holzinger, columns = 7:15, mapping = aes(alpha = .1, colour = color))
```


## Model specification

Your latent variables' names are arbitrary. Your manifest variables should be column names from your dataset. 

```{r}
model_def_visual_IQ <- 'textspeed =~ x1 + x2 + x3 +x7 + x8 + x9'
```


Model quality is assessed based on degrees of freedom.

Degrees of Freedom (df) are determined by the number of manifest variables and estimated values, where:


* df = Possible Values - Estimated Values

* Possible Values = Manifest Variables * (Manifest Variables + 1) / 2


**Remember**: 

* Include at least three manifest variables

* Create models with df > 0


## Fit the model

```{r}
model_fit_visual_IQ <- cfa(model = model_def_visual_IQ,
                           data  = holzinger)
```


## Interpret the model

```{r}
summary(model_fit_visual_IQ)
```


It might be hard to interpret estimates as they are expressed in the scale of the data. 

We can standardize the solution to use z-scores:


```{r}
summary(model_fit_visual_IQ, standardized = TRUE)
```


This produces a new column `Std.all`. Values close to 1 indicate a strong relationship to the latent variable. 

### Explore model fit

```{r}
summary(model_fit_visual_IQ, fit.measures = TRUE)
```

It seems that the fit is not very good. $TLI$ and $CFI$ are much below $.9 - .95$ range, which is a customary threshold for good model fit. $RMSEA > .1$, which is another indicator of very average fit. 


